Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-10-02 paper.tex
improve cycle staggering bad example plot

One difference between the fleet and individual facility modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot C$
of power for that time step where $C$ is the power capacity of a single
reactor of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is
short a half batch of fuel, then it will produce no power ($0 \cdot C$) that
time step.  Time step duration can also have this effect.  For both fleet and
individual reactor modeling, power goes offline in full time step increments.
For example, if a 3-batch fleet based reactor is short $\frac{1}{2}$ batch of
fuel with a 3 month time step, then it will produce $\frac{2.5}{3} \cdot 3
\text{mo} \cdot C $ of energy.  In contrast, a monthly time step might allow
the same reactor to produce $[\frac{5}{6} \cdot 1 \text{mo} + 1 \cdot 2
\text{mo}] \cdot C $ of energy if the fuel shortage actually ended after only
1 month.  Quantifying this effect directly is difficult.  One way of observing
it is to compare fleet reactor scenarios that have different time step
durations (i.e. cases MF and QF).  Because the fleet reactors have perfect fuel
sharing (described below) and no outage noise, differences observed will
likely be due to quantized shutdown.

Individually modeled reactors each have their own refueling cycles.  Depending
on how the refueling cycles and outages are staggered, power production can
vary significantly.  For example, if all reactors refuel at the same time,
then all reactors are offline at the same time.  This is referred to as the
\emph{cycle staggering effect} and does not occur in fleet-based reactor
models.  An example of poor cycle staggering is shown in Figure
\ref{fig:sync-cycle}.  This effect can be naturally observed as the
fluctuations in the difference between deployed power capacity and generated
power.

Poor reactor cycle staggering can also cause spikes of material supply and
demand that can lead to suboptimal resource utilization.  Consider a fuel
fabrication facility that receives requests for new batches for every reactor
all at once.  The reactors draw out inventory from the fuel fabrication
facility together in a large quantity of orders.  If the fabrication facility
does not have enough material on hand, some number of reactors will need to
wait until enough fuel can be pumped through fabrication.  Avoiding such
constraints would require the fabrication facility to maintain suboptimally
large on-hand inventories as a contingency.  Even with infinite material
availability, poor staggering can result in supply constraints caused by
finite facility throughput. Even if the fuel fabrication facility above has
infinite material supply that it can keep on hand, it might not be able to
fabricate fuel for all reactor requests all at once even though it has
sufficient capacity if they request fuel uniformly over time.

\begin{figure}[h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/sync-cycle.eps}
    \caption[The cycle staggering effect]{
        In this scenario, reactors are deployed annualy with a refueling cycle
        length (including outage) of 18 months.  So all reactors deployed
        every 3rd year refuel together.  All initial 100 LWRs start out with
        their cycles synced as well.  While it may look like separate curves,
        this plot shows only a single time series.
    }
    \label{fig:sync-cycle}
\end{figure}

>>>>

One difference between the fleet and individual facility modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot C$
of power for that time step where $C$ is the power capacity of a single
reactor of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is
short a half batch of fuel, then it will produce no power ($0 \cdot C$) that
time step.  Time step duration can also have this effect.  For both fleet and
individual reactor modeling, power goes offline in full time step increments.
For example, if a 3-batch fleet based reactor is short $\frac{1}{2}$ batch of
fuel with a 3 month time step, then it will produce $\frac{2.5}{3} \cdot 3
\text{mo} \cdot C $ of energy.  In contrast, a monthly time step might allow
the same reactor to produce $[\frac{5}{6} \cdot 1 \text{mo} + 1 \cdot 2
\text{mo}] \cdot C $ of energy if the fuel shortage actually ended after only
1 month.  Quantifying this effect directly is difficult.  One way of observing
it is to compare fleet reactor scenarios that have different time step
durations (i.e. cases MF and QF).  Because the fleet reactors have perfect fuel
sharing (described below) and no noise from refueling outages, differences observed will
likely be due to quantized shutdown.

Individually modeled reactors each have their own refueling cycles.  Depending
on how the refueling cycles and outages are staggered, power production can
vary significantly.  For example, if all reactors refuel at the same time,
then all reactors are offline at the same time.  This is referred to as the
\emph{cycle staggering effect} and does not occur in fleet-based reactor
models.  An example of poor cycle staggering is shown in Figure
\ref{fig:sync-cycle}.  This effect can be naturally observed as the
fluctuations in the difference between deployed power capacity and generated
power.  The figure shows power jumping above and below the installed net
capacity (includes capacity factor) from time step to time step depending on
how many reactors are online/offline together.

Poor reactor cycle staggering can also cause spikes of material supply and
demand that can lead to suboptimal resource utilization.  Consider a fuel
fabrication facility that receives requests for new batches for every reactor
all at once.  The reactors draw out inventory from the fuel fabrication
facility together in a large quantity of orders.  If the fabrication facility
does not have enough material on hand, some number of reactors will need to
wait until enough fuel can be fabricated.  Avoiding such
constraints would require the fabrication facility to maintain suboptimally
large on-hand inventories as a contingency.  Even with infinite material
availability, poor staggering can result in supply constraints caused by
finite facility throughput. Even if the fuel fabrication facility above has
infinite material supply that it can keep on hand, it might not be able to
fabricate fuel for all reactor requests all at once even though it has
sufficient capacity if they request fuel uniformly over time.

\begin{figure}[h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/sync-cycle.eps}
    \caption[The cycle staggering effect]{
        In this scenario, reactors are deployed annualy with a refueling cycle
        length (including outage) of 18 months.  So all reactors deployed
        every 3rd year refuel together.  All initial 100 LWRs start out with
        their cycles synced as well.  While it may look like separate curves,
        generated power curve (purple) represents only a single time series with many discontinuities.
    }
    \label{fig:sync-cycle}
\end{figure}
